---
title: '{rcahelpr}'
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## ❓ What and Why is This?
In short, a tool to support the Resident Corrections Analyst program through bespoke functions written in R and data sets for learning purposes. Interested? You can install is using R:

```{r, eval=FALSE}
remotes::install_github("cllghn/rcahelpr")
```


## 🔎 Examples

### 📚 Codebooks

Codebooks are often to provide a comprehensive guide to the variables and coding schemes in a data set, ensuring consistent and accurate interpretation of data. They serve as a reference tool to facilitate understanding and analysis of complex data structures by researchers and analysts. To support this crucial functionality, we include a codebook function into this package:

```{r}
# Load the library
library(rcahelpr)

# Create a test data set
test <- data.frame(
  "person" = c("chris", "maeve", "joseph", "brooks"),
  "org" = c("csg", "wdoc", "ccjbh", "asu"),
  "years_in_org" = c(1, 0.3, 0.2, NA),
  "role" = factor(c("mentor", "rca", "rca", "mentor"))
)

# Create codebook
make_codebook(input_df = test, return_df = FALSE, escape = FALSE)
```

Should you want to add information to describe the variables described in the codebook, you can do so by left joining and additional data set:

```{r}
# Set a secondary data.frame describing the variables in your original data set
more <- data.frame(
  "vars" = c("person", "org", "years_in_org", "role"),
  "description" = rep("Interesting details about my variable.", 4),
  "origin" = rep("Detailed notes on where the data came from.", 4),
  "notes" = rep("Yet more useful information", 4)
)

# Create codebook
make_codebook(input_df = test, return_df = FALSE, escape = FALSE,
              extra_vars = more, extra_key = "vars")
```

### 📊 Data

Some key data sets, for learning or general use have been included in the library. Access them using the `::` accessor:

```{r}
str(rcahelpr::hpsa_primarycare)
```

This means that we can use these data for a range of purposes, such as pairing them with other libraries and analyzing it:

```{r, fig.height=10}
# Load the graphing library ggplot2 and data management library dplyr
library(dplyr)
library(ggplot2)

# Wrangle some data to identify the average Health Professional Shortage Area 
# (HPSA) score in a given county:
demo <- rcahelpr::hpsa_primarycare %>%
  group_by(County_Equivalent_Name) %>%
  summarize(mean_hpsa_score = mean(HPSA_Score),
            total_hpsa_population = sum(HPSA_Designation_Population))

# Make a beautiful graph
ggplot(data = demo) +
  geom_point(aes(x = total_hpsa_population , y = County_Equivalent_Name,
                 color = mean_hpsa_score)) +
  geom_segment(aes(x = 0 , xend = total_hpsa_population ,
                   y = County_Equivalent_Name, yend = County_Equivalent_Name,
                   color = mean_hpsa_score)) +
  theme_minimal() +
  labs(title = "Affected Populations in HPSA by County",
       subtitle = "HPSA scores determine priorities for the assignment of clinitians",
       caption = "Data from CCHS") +
  xlab("Total Population in all County HPSAs") +
  ylab("") +
  scale_color_gradient2(low="#F5F5DC", mid = "#FFA500", high="#8B0000", 
                        midpoint = mean(demo$mean_hpsa_score),
                        name = "Average County HPSA Score")


```

