---
title: '📚 Codebooks'
output: html_document
always_allow_html: true
---

Codebooks are often to provide a comprehensive guide to the variables and coding schemes in a data set, ensuring consistent and accurate interpretation of data. They serve as a reference tool to facilitate understanding and analysis of complex data structures by researchers and analysts. 

To support this crucial functionality, we include a codebook function into this package:

```{r}
# Load the library
library(rcahelpr)

# Create a test data set
test <- data.frame(
  "person" = c("chris", "maeve", "joseph", "brooks"),
  "org" = c("csg", "wdoc", "ccjbh", "asu"),
  "years_in_org" = c(1, 0.3, 0.2, NA),
  "role" = as.factor(c("mentor", "rca", "rca", "mentor")),
  "date" = as.Date(c("2020-01-01", "2020-01-01", NA, "2020-01-02"))
)

# Create codebook
make_codebook(input_df = test, return_df = FALSE, escape = FALSE)
```

Should you want to add information to describe the variables described in the codebook, you can do so by left joining and additional data set:

```{r}
# Set a secondary data.frame describing the variables in your original data set
more <- data.frame(
  "vars" = c("person", "org", "years_in_org", "role"),
  "description" = rep("Interesting details about my variable.", 4),
  "origin" = rep("Detailed notes on where the data came from.", 4),
  "notes" = rep("Yet more useful information", 4)
)

# Create codebook
make_codebook(input_df = test, return_df = FALSE, escape = FALSE,
              extra_vars = more, extra_key = "vars")
```
